---
title: "Gottumukkula_Sriya_A02_CodingBasics.Rmd"
author: "Sriya Gottumukkula"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons/labs in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Canvas.

## Basics, Part 1

1.  Generate a sequence of numbers from one to 55, increasing by fives. Assign this sequence a name.

```{r}
data1 <- seq(1,55,5) #creating a sequence 
```


2.  Compute the mean and median of this sequence.

```{r}
meandata1 <- mean(data1) #calculating the mean 

mediandata1 <-median(data1) # calculating the median 
```


3.  Ask R to determine whether the mean is greater than the median 

```{r}
comparisondata1 <-(meandata1>mediandata1) # seeing if mean is greater than median 
```


4.  Insert comments in your code to describe what you are doing.

```{r basics 1}
#1. creating a sequence 

#2. calculating mean and median 

#3. seeing if mean is greater than median 

```

## Basics, Part 2

5.  Create three vectors, each with four components, consisting of (a) student names, (b) test scores, and (c) whether they are on scholarship or not (TRUE or FALSE). 

```{r}
names <- c("sriya","priya", "niya", "giya") #character vector 
scores <- c(10,20,30,40) #numerical vector 
scholarship <- c("yes","yes","no","no") #character vector

```


6.  Label each vector with a comment on what type of vector it is.

```{r}
#1 Character Vector 
#2 Numerical Vector
#3 Character Vector 
```


7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

```{r}
dataframe <- data.frame("names","scores","scholarship")
```


8.  Label the columns of your data frame with informative titles.

```{r basics 2}
colnames(dataframe) <- c("names","scores","scholarship")

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: this data frame is different from a matrix because a data frame can hold multiple data types (characters, numbers) in different columns, while a matrix can only contain a single data type across all its elements.

10. Create a function with one input. In this function, use `if`...`else` to evaluate the value of the input: if it is greater than 50, print the word "Pass"; otherwise print the word "Fail". 

```{r}
code1 <- function(x) {
  if(x > 50) {
   (x <- "pass")
  } 
  else {
    (x <- "fail")
  }
}


```


11. Create a second function that does the exact same thing as the previous one but uses `ifelse()` instead if `if`...`else `. 

```{r}
code2 <- function(x){
  ifelse(x>50, "pass", "fail") 
  
}

```


12. Run both functions using the value 52.5 as the input
```{r}
output1 <- code1(52.5)
output2 <- code2(52.5)
```

13. Run both functions using the **vector** of student test scores you created as the input. (Only one will work properly...)

```{r}
#output3 <- code1(scores)
output4 <- code2(scores)
```


```{r basics 3}
#10. Create a function using if...else

code3 <- function(x) {
  if(x < 4) {
    x*3
  } 
  else {
    x*2
  }
}

#11. Create a function using ifelse()

code4 <- function(x){
  ifelse(x<4, x*3, x*2) 
  
}

#12a. Run the first function with the value 52.5

output5 <- code3(52.5)

#12b. Run the second function with the value 52.5

output6 <- code4(52.5)

#13a. Run the first function with the vector of test scores 

#output7 <- code3(scores)

#13b. Run the second function with the vector of test scores

output8 <- code4(scores)

```

14. QUESTION: Which option of `if`...`else` vs. `ifelse` worked? Why? (Hint: search the web for "R vectorization")



> Answer: only ifelse worked as vectors only work with ifelse() and not if and else because if statements in R are designed to handle a single logical condition, not multiple values. When a vector is used in an if statement, R only checks the first element and ignores the rest, leading to incorrect results or warnings. In contrast, ifelse() is vectorized, meaning it applies the condition element-wise to each value in the vector, returning a result for each one


**NOTE** Before knitting, you'll need to comment out the call to the function in Q13 that does not work. (A document can't knit if the code it contains causes an error!)